Hyperbolic SMC: a Variational Proposal for Sequential Monte Carlo Phylogenetic Inference

### 2024-07-29 paper/paper.tex
Write abstract

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

>>>>

\begin{abstract}
  Variational combinatorial sequential Monte Carlo (VCSMC) provides a robust and flexible framework for learning distributions over phylogenetic state space. In the original VCSMC implementation, learnable branch length parameters are linked to each coalescent event. We propose a new variational proposal distribution in which the tips of the phylogenetic tree are embedded in a hyperbolic latent space that captures the evolutionary distance between biological sequences. We show that our proposal parameters can be efficiently optimized to maximize the evidence lower bound (ELBO) using stochastic gradient descent. Our method, Hyperbolic SMC, produces likelihood estimates that are competitive with state-of-the-art combinatorial sequential Monte Carlo (CSMC) methods. Hyperbolic SMC produces tight likelihood estimates on even very large datasets, requires no preliminary tree search, and is highly parallelized to take advantage of modern GPUs.
\end{abstract}
